---
title: "Penguin Flipper Lengths ANOVA"
author: "Renee Albrecht"
date: "November 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read in the penguins.csv data

Packages: tidyverse, car, onewaytests

```{r}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")

View(penguins)


```

Our question: Is there a significnat difference in mean flipper lengths for female Chinstrap, Gentoo, and Adelie Penguins near Palmer Archipelago? 

Maybe I can use a one way ANOVA...
Assumptions: 
- Independence obervations 
- Equal variance (Levene's Test)
- Normality (histogram/qqplot)

```{r}

counts <- with(penguins, table(species, sex))

penguins2 <- penguins %>%
  mutate(common_name = 
  case_when(
    species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap", 
    species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
    species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie")) %>%
  select(-species) %>%
  filter(sex == "FEMALE")

View(penguins2)
```

Exploratory graphs and Levene's test for equal variances 

```{r}
hists <- ggplot(penguins2, aes(x = flipper_length)) +
  geom_histogram(aes(fill = common_name)) +
  facet_wrap(~ common_name, scale = "free")

hists
  
  
qqs <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq(aes(color = common_name)) +
  facet_wrap(~ common_name, scale = "free")

qqs
  
  
  
## Based on qq plots and histograms, observations for each group appear normally distributed. Also, n > 30 for all groups, so means will be normally distributed by CLT. 


```

Evaluate variances and perform Levene's Test for equal variances 

```{r}

#If the largest sample variance is <4x greater than the smallest sample variance, then these are "close enough" to consider variances equal 

variances <- penguins2 %>%
  group_by(common_name) %>%
  summarize(
    variance = var(flipper_length)
  )

variances #The largest variance (33.1) is definitely not more than 4x greater than the smallest variance (15.2)

# Levene's Test (analagous test to F test if there are more than 2 groups)
# leveneTest function in the 'car' package 

levene_penguin <- leveneTest(flipper_length ~ common_name, data = penguins2)
levene_penguin

# H0: No difference in variances (variances are equal)
# HA: Variances are NOT equal 

# There is no significant difference in variances across species 

```

One-way ANOVA for penguin flipper lengths by species (common_name)

```{r}

# H0: Means across all groups are equal 
# HA: At least two group means differ significantly 


penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)
summary(penguin_aov)

#Post-hoc test by Tukey's HSD
#H0: No significant difference in pairwise means 
#H0: There IS a significant difference in pairwise means 

penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph

# aov.test function in the 'onewaytests' package DOES allow for inline references 

```


Mean flipper lengths of female penguins differed significantly for all species studied (one-way ANOVA, d(df) = statistic value, p < 0.001, $\alpha$ = 0.05 with post-hoc Tukey's HSD, $\alpha$ = 0.05). 




